Penalized linear mixed models for structured genetic data

### 2021-03-14 main.tex
Edits to 5 and 6

@@

\pb{uh, what? This formula does NOT require sign to be correct.  What's right, the formula or the text?}

Selection accuracy was evaluated using selection precision. All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly. Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/genomewide_prc_p1=100.png}
\caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_prc}
\end{figure}

PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selection accuracy, particularly for the first $\approx$ 100 SNPs each method allows into the model. For the most part, PC-lasso selects SNPs outperforms LMM-lasso in this setting as well. All methods struggle as the model becomes less sparse and the number of selected SNPs approaches $n$.

%------------------------------%
\subsection{\anna{PC-lasso and LMM-lasso increase genome-wide estimation accuracy}}
%------------------------------%

@@

In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso. This bias is most dramatic for the first $\approx$ 100 SNPs selected by each method. Both PC-lasso and LMM-lasso produce more accurate estimates than the ordinary lasso for any number of selected SNPs.

@@

%------------------------------%
\subsection{\anna{PC-lasso and LMM-lasso reduce selection bias in genome-wide data}}
%------------------------------%

@@

Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected for a selected SNPs with the first principal component (PC1) of the Empirical data for each number of incorrectly selected SNPs. The corresponding shaded regions indicate the SE about the mean. \anna{Do I need to explain why we are using correlation with the first PC as a proxy for MAF difference or is this obvious?} The absolute correlation with PC1 is used as a measure of how population-informative the selected SNPs are given the complex Empirical data structure compared to that in Section~\ref{sec:sel_bias}.

@@

The ordinary lasso is clearly most susceptible to selecting non-causal SNPs due to their association with population structure. This corresponds to the simulation results shown in Section~\ref{sec:sel_bias}, though in Figure \ref{fig:gw_pc1_corr} we more plainly see that this bias persists regardless of the number of variables allowed into the model. Likewise, PC-lasso and LMM-lasso both effectively control this type of selection bias. PC-lasso's effective control of bias in this setting is to be expected, given that we are assessing the population-informative nature of incorrectly selected SNPs via correlation with PC1, which is explicitly adjusted for in the PC-lasso method.

@@

\anna{The results of a genome-wide analysis of Empirical data agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso. We note that because the Empirical data consists of three distinct racial groups, its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates predominantly coarse population structure. Given our results in Section~\ref{sec:sim-fine-coarse}, we would expect LMM-lasso to perform even better in the analysis of more finely structured genome-wide data.}

>>>>

\pb{How do you feel about combining Sections 5.1 + 5.2, and Figures 5 + 6?  Each section seems really short to me.}

@@

\pb{I'm confused -- this formula does NOT require sign to be correct.  What's right, the formula or the text?}

Selection accuracy was evaluated using selection precision. All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly. Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/genomewide_prc_p1=100.png}
\caption{\pb{I feel like we should stop this at 200.  The 100-400 range dominates this plot and what I get from this figure is: all the methods are equally bad.} Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_prc}
\end{figure}

PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly for the first $\approx$ 100 SNPs each method allows into the model. For the most part, PC-lasso outperforms LMM-lasso in this setting as well. None of the method are able to accurately select more than about 100 SNPs.  However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first \pb{XXX} SNPs selected were all false positives.

%------------------------------%
\subsection{PC-lasso and LMM-lasso increase genome-wide estimation accuracy}
%------------------------------%

@@

In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso. This bias is most dramatic for the first $\approx$ 100 SNPs selected by each method. For any number of selected SNPs, however, both PC-lasso and LMM-lasso produce much more accurate estimates than the ordinary lasso.

@@

%------------------------------%
\subsection{PC-lasso and LMM-lasso reduce selection bias in genome-wide data}
%------------------------------%

@@

Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected with the first principal component (PC1). The first principal component does not capture the full complexity of the population structure, but it useful in terms of quantifying the extent to which a method is selecting SNPs that merely reflect the population structure and hence are at greater risk of being false positives due to environmental confounding. The shaded regions indicate the SE about the mean.

@@

The ordinary lasso is clearly most susceptible to selecting non-causal SNPs due to their association with population structure. This corresponds to the simulation results shown in Section~\ref{sec:sel_bias}, though in Figure \ref{fig:gw_pc1_corr} we more plainly see that this bias persists regardless of the number of variables allowed into the model. Likewise, PC-lasso and LMM-lasso both effectively control this type of selection bias. PC-lasso's effective control of bias in this setting is to be expected, given that we are explicitly adjusting for PC1 in the analysis, but we see here that LMM-lasso accomplishes the same thing in a more indirect manner.

@@

\pb{I did change this paragraph a bit; I feel that we need to acknowledge and be honest about the fact that PC-lasso outperformed LMM-lasso here.}
The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso. However, PC-lasso performed somewhat better than LMM-lasso in this case study, in contrast to most of the simulation experiments. This is perhaps to be expected as the sample consisted of three distinct racial groups with little admixture, and its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates a rather coarse population structure. Thus, its structure is well-suited for PC-lasso. This finding suggests that while LMM-lasso may be more robust to complex structures, PC-lasso remains a strong competitor when the population structure is coarse.
